### 2022-01-16 Check_Data.Rmd
First update of all files

---
title: "Check_Data"
author: "Amalan"
date: "21/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.width = 7,fig.height = 7,warning = FALSE,message = FALSE)
library(readr)
library(here)
library(dplyr)
library(corrr)
library(corrplot)
library(GGally)
```

# Model 1 Testing

```{r Model 1 Testing}
load(here("Identical_r0","Generate_Big_Data","Model_1","No_Correlated_Covariate.RData"))
Original_Data<-as.data.frame(Simulated_Data$All_Data$Assumed_Model_3)
no_of_Variables<-2

colnames(Original_Data)<-c("Y","X0",paste0("X",1:no_of_Variables),
                           paste0("X",1:no_of_Variables,"^2"))

glm(Y~.-1,data=as.data.frame(Original_Data),family="binomial")->Model_1
stepAIC(Model_1,direction = "backward")
```

# Model 2 Testing

```{r Model 2 Testing}
load(here("Identical_r0","Generate_Big_Data","Model_2","No_Correlated_Covariate.RData"))
Original_Data<-as.data.frame(Simulated_Data$All_Data$Assumed_Model_3)
no_of_Variables<-2
colnames(Original_Data)<-c("Y","X0",paste0("X",1:no_of_Variables),
                           paste0("X",1:no_of_Variables,"^2"))
glm(Y~.-1,data=as.data.frame(Original_Data),family="binomial")->Model_2
stepAIC(Model_2,direction = "backward")
```

# Model 3 Testing

```{r Model 3 Testing}
load(here("Identical_r0","Generate_Big_Data","Model_3","No_Correlated_Covariate.RData"))
Original_Data<-as.data.frame(Simulated_Data$All_Data$Assumed_Model_3)
no_of_Variables<-2
colnames(Original_Data)<-c("Y","X0",paste0("X",1:no_of_Variables),
                           paste0("X",1:no_of_Variables,"^2"))
glm(Y~.-1,data=as.data.frame(Original_Data),family="binomial")->Model_3
stepAIC(Model_3,direction = "backward")
```

# Model 4 Testing

```{r Model 4 Testing}
load(here("Identical_r0","Generate_Big_Data","Model_4","No_Correlated_Covariate.RData"))
Original_Data<-as.data.frame(Simulated_Data$All_Data$Real_Model)
no_of_Variables<-2
colnames(Original_Data)<-c("Y","X0",paste0("X",1:no_of_Variables),
                           paste0("X",1:no_of_Variables,"^2"))
glm(Y~.-1,data=as.data.frame(Original_Data),family="binomial")->Model_4
stepAIC(Model_4,direction = "backward")
```