### 2023-02-10 Identical_r0/Rmarkdown/OSMAC_Method.Rmd

---
params:
  Model_Path:
title: "OSMAC_Method; Model: `r params$Model_Path`; "
author: "M.Amalan"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,fig.height = 9,fig.width = 9,message = FALSE,warning=FALSE)
library(ggplot2)
library(here)
library(dplyr)
library(tidyr)
library(gganimate)
library(LaplacesDemon)
library(linguisticsdown)
library(gganimate)
library(viridis)
library(compiler)
library(data.table)
library(ggpubr)

# Cordeiro Error ----
Cordeiro<-function(XData,With_bias)
{
  p <- as.vector(invlogit(XData%*%as.vector(With_bias)))
  W <- diag(p*(1-p))
  inverse_term <- solve(t(XData)%*%W%*%XData)

  Term1 <- inverse_term%*%t(XData)%*%W
  Term2 <- diag(diag(XData%*%(inverse_term)%*%t(XData))) %*% as.matrix(p-0.5)

  bias <- as.vector(Term1%*%Term2)
  return(bias)
}

Cordeiro<-cmpfun(Cordeiro)

FullDataPath<-here("Identical_r0","Generate_Big_Data",params$Model_Path,"No_Correlated_Covariate.RData")
load(here(FullDataPath))
Full_model<-glm(Y~.-1,data = as.data.frame(Simulated_Data[[2]]$Real_Model),family = binomial)
Real_Vline<-data.frame("Theta"=c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1]))),
                       "Values"=Simulated_Data$Basic$Theta)
Estimated_Vline<-data.frame("Theta"=c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1]))),
                            "Values"=Full_model$coefficients)
Bias_Vline<-data.frame("Theta"=c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1]))),"Values"=0)
```

# OSMAC ;  Response Model : `r params$Model_Path`

## Real Model Data

```{r Load Data}
Path<-here("Identical_r0","Outputs",params$Model_Path,"OSMAC","Real_Model/")
load(here(Path,"OSMAC_output.RData"))

Est_Parameter_RM<-rbind.data.frame(Results_OSMAC$mMSE_Output,
                                   Results_OSMAC$mVc_Output)

Est_Bias_RM<-rbind.data.frame(Bias_OSMAC$mMSE_Output,
                              Bias_OSMAC$mVc_Output)

Utility_RM<-rbind.data.frame(Utility_OSMAC$mMSE_Output,
                             Utility_OSMAC$mVc_Output)

colnames(Est_Parameter_RM)<-colnames(Est_Bias_RM)<-c("Type","Subsample_Size",
                                                     paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))
colnames(Utility_RM)<-c("Type","Subsample_Size","A_optimality","D_optimality")

remove(Results_OSMAC,Bias_OSMAC,Utility_OSMAC)
```

## Assumed Model Data

### Old Probabilities

```{r Load Data Assumed Models}
Path<-here("Identical_r0","Outputs",params$Model_Path,"OSMAC","Assumed_Model/")
# Assumed Model 1
load(here(Path,"OSMAC_output_Old_1.RData"))

Est_Parameter_AM_Old_1<-rbind.data.frame(Results_OSMAC$mMSE_Output,
                                         Results_OSMAC$mVc_Output)

Est_Bias_AM_Old_1<-rbind.data.frame(Bias_OSMAC$mMSE_Output,
                                    Bias_OSMAC$mVc_Output)

Utility_AM_Old_1<-rbind.data.frame(Utility_OSMAC$mMSE_Output,
                                   Utility_OSMAC$mVc_Output)

colnames(Est_Parameter_AM_Old_1)<-colnames(Est_Bias_AM_Old_1)<-c("Type","Subsample_Size",
                                                                 paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))
colnames(Utility_AM_Old_1)<-c("Type","Subsample_Size","A_optimality","D_optimality")

remove(Results_OSMAC,Bias_OSMAC,Utility_OSMAC)

# Assumed Model 2
load(here(Path,"OSMAC_output_Old_2.RData"))

Est_Parameter_AM_Old_2<-rbind.data.frame(Results_OSMAC$mMSE_Output,
                                         Results_OSMAC$mVc_Output)

Est_Bias_AM_Old_2<-rbind.data.frame(Bias_OSMAC$mMSE_Output,
                                    Bias_OSMAC$mVc_Output)

Utility_AM_Old_2<-rbind.data.frame(Utility_OSMAC$mMSE_Output,
                                   Utility_OSMAC$mVc_Output)

colnames(Est_Parameter_AM_Old_2)<-colnames(Est_Bias_AM_Old_2)<-c("Type","Subsample_Size",
                                                                 paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))
colnames(Utility_AM_Old_2)<-c("Type","Subsample_Size","A_optimality","D_optimality")

remove(Results_OSMAC,Bias_OSMAC,Utility_OSMAC)

# Assumed Model 3

load(here(Path,"OSMAC_output_Old_3.RData"))

Est_Parameter_AM_Old_3<-rbind.data.frame(Results_OSMAC$mMSE_Output,
                                         Results_OSMAC$mVc_Output)

Est_Bias_AM_Old_3<-rbind.data.frame(Bias_OSMAC$mMSE_Output,
                                    Bias_OSMAC$mVc_Output)

Utility_AM_Old_3<-rbind.data.frame(Utility_OSMAC$mMSE_Output,
                                   Utility_OSMAC$mVc_Output)

colnames(Est_Parameter_AM_Old_3)<-colnames(Est_Bias_AM_Old_3)<-c("Type","Subsample_Size",
                                                                 paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))
colnames(Utility_AM_Old_3)<-c("Type","Subsample_Size","A_optimality","D_optimality")

remove(Results_OSMAC,Bias_OSMAC,Utility_OSMAC)
```

### New Probabilities

```{r Load Data Assumed Models 2}
Path<-here("Identical_r0","Outputs",params$Model_Path,"OSMAC","Assumed_Model/")
# Assumed Model 1
load(here(Path,"OSMAC_output_New_1.RData"))

Est_Parameter_AM_New_1<-rbind.data.frame(Results_OSMAC$mMSE_Output,
                                         Results_OSMAC$mVc_Output)

Est_Bias_AM_New_1<-rbind.data.frame(Bias_OSMAC$mMSE_Output,
                                    Bias_OSMAC$mVc_Output)

Utility_AM_New_1<-rbind.data.frame(Utility_OSMAC$mMSE_Output,
                                   Utility_OSMAC$mVc_Output)

colnames(Est_Parameter_AM_New_1)<-colnames(Est_Bias_AM_New_1)<-c("Type","Subsample_Size",
                                                                 paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))
colnames(Utility_AM_New_1)<-c("Type","Subsample_Size","A_optimality","D_optimality")

remove(Results_OSMAC,Bias_OSMAC,Utility_OSMAC)

# Assumed Model 2
load(here(Path,"OSMAC_output_New_2.RData"))

Est_Parameter_AM_New_2<-rbind.data.frame(Results_OSMAC$mMSE_Output,
                                         Results_OSMAC$mVc_Output)

Est_Bias_AM_New_2<-rbind.data.frame(Bias_OSMAC$mMSE_Output,
                                    Bias_OSMAC$mVc_Output)

Utility_AM_New_2<-rbind.data.frame(Utility_OSMAC$mMSE_Output,
                                   Utility_OSMAC$mVc_Output)

colnames(Est_Parameter_AM_New_2)<-colnames(Est_Bias_AM_New_2)<-c("Type","Subsample_Size",
                                                                 paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))
colnames(Utility_AM_New_2)<-c("Type","Subsample_Size","A_optimality","D_optimality")

remove(Results_OSMAC,Bias_OSMAC,Utility_OSMAC)

# Assumed Model 3
load(here(Path,"OSMAC_output_New_3.RData"))

Est_Parameter_AM_New_3<-rbind.data.frame(Results_OSMAC$mMSE_Output,
                                         Results_OSMAC$mVc_Output)

Est_Bias_AM_New_3<-rbind.data.frame(Bias_OSMAC$mMSE_Output,
                                    Bias_OSMAC$mVc_Output)

Utility_AM_New_3<-rbind.data.frame(Utility_OSMAC$mMSE_Output,
                                   Utility_OSMAC$mVc_Output)

colnames(Est_Parameter_AM_New_3)<-colnames(Est_Bias_AM_New_3)<-c("Type","Subsample_Size",
                                                                 paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))
colnames(Utility_AM_New_3)<-c("Type","Subsample_Size","A_optimality","D_optimality")

remove(Results_OSMAC,Bias_OSMAC,Utility_OSMAC)
```

## Model Free

```{r Load Data Model Free}
Path<-here("Identical_r0","Outputs",params$Model_Path,"OSMAC","ModelFree/")
load(here(Path,"OSMAC_output.RData"))

Est_Parameter_MF<-rbind.data.frame(Results_OSMAC$mMSE_Output,
                                   Results_OSMAC$mVc_Output)

Est_Bias_MF<-rbind.data.frame(Bias_OSMAC$mMSE_Output,
                              Bias_OSMAC$mVc_Output)

Utility_MF<-rbind.data.frame(Utility_OSMAC$mMSE_Output,
                             Utility_OSMAC$mVc_Output)

colnames(Est_Parameter_MF)<-colnames(Est_Bias_MF)<-c("Type","Subsample_Size",
                                                             paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))
colnames(Utility_MF)<-c("Type","Subsample_Size","A_optimality","D_optimality")

remove(Results_OSMAC,Bias_OSMAC,Utility_OSMAC)
```

```{r Combine all datasets}
Est_Parameter<-rbind(cbind(Model="Real",Est_Parameter_RM),
                     cbind(Model="Assumed Old 1",Est_Parameter_AM_Old_1),
                     cbind(Model="Assumed Old 2",Est_Parameter_AM_Old_2),
                     cbind(Model="Assumed Old 3",Est_Parameter_AM_Old_3),
                     cbind(Model="Assumed New 1",Est_Parameter_AM_New_1),
                     cbind(Model="Assumed New 2",Est_Parameter_AM_New_2),
                     cbind(Model="Assumed New 3",Est_Parameter_AM_New_3),
                     cbind(Model="Model Free",Est_Parameter_MF))

remove(Est_Parameter_RM,Est_Parameter_MF,
       Est_Parameter_AM_Old_1,Est_Parameter_AM_Old_2,Est_Parameter_AM_Old_3,
       Est_Parameter_AM_New_1,Est_Parameter_AM_New_2,Est_Parameter_AM_New_3)

Est_Bias<-rbind(cbind(Model="Real",Est_Bias_RM),
                cbind(Model="Assumed Old 1",Est_Bias_AM_Old_1),
                cbind(Model="Assumed Old 2",Est_Bias_AM_Old_2),
                cbind(Model="Assumed Old 3",Est_Bias_AM_Old_3),
                cbind(Model="Assumed New 1",Est_Bias_AM_New_1),
                cbind(Model="Assumed New 2",Est_Bias_AM_New_2),
                cbind(Model="Assumed New 3",Est_Bias_AM_New_3),
                cbind(Model="Model Free",Est_Bias_MF))

remove(Est_Bias_RM,Est_Bias_MF,
       Est_Bias_AM_Old_1,Est_Bias_AM_Old_2,Est_Bias_AM_Old_3,
       Est_Bias_AM_New_1,Est_Bias_AM_New_2,Est_Bias_AM_New_3)

Utility<-rbind(cbind(Model="Real",Utility_RM),
               cbind(Model="Assumed Old 1",Utility_AM_Old_1),
               cbind(Model="Assumed Old 2",Utility_AM_Old_2),
               cbind(Model="Assumed Old 3",Utility_AM_Old_3),
               cbind(Model="Assumed New 1",Utility_AM_New_1),
               cbind(Model="Assumed New 2",Utility_AM_New_2),
               cbind(Model="Assumed New 3",Utility_AM_New_3),
               cbind(Model="Model Free",Utility_MF))

remove(Utility_RM,Utility_MF,
       Utility_AM_Old_1,Utility_AM_Old_2,Utility_AM_Old_3,
       Utility_AM_New_1,Utility_AM_New_2,Utility_AM_New_3)
```

## Parameters

```{r Animated Plot for Parameter}
Est_Parameter %>%
  pivot_longer(starts_with("Theta"), names_to = "Theta", values_to = "Estimated") %>%
ggplot(., aes(x = Estimated, group = as.factor(Model), frame=Subsample_Size,fill=factor(Model))) + 
  geom_histogram(binwidth = 0.02,alpha=0.8)+
  facet_grid(Type~Theta,scales = "free_x",shrink = TRUE)+
  geom_vline(data= Estimated_Vline,aes(xintercept = Values,color="red"),linetype='dashed',size=1)+
  scale_color_discrete(name = "Parameter values", labels = c("Estimated"))+
  labs(title=paste0("Real Model :",params$Model_Path," Data : Estimated Theta"),
      subtitle = "Sub-sample Size: {current_frame}")+
  transition_manual(Subsample_Size)+
  theme_light()+ylab("Frequency")+xlab("Estimates")->para_1

anim_save(filename="para_1.gif",animation = para_1,
          path=here("Identical_r0","htmloutputs",params$Model_Path,"OSMAC","Plots"))

include_graphics2(here("Identical_r0","htmloutputs",params$Model_Path,"OSMAC","Plots","para_1.gif"))
```

## Bias

```{r Animated Plots for Bias RM}
Est_Bias %>%
  pivot_longer(starts_with("Theta"), names_to = "Theta", values_to = "Estimated") %>%
ggplot(., aes(x = Estimated, group = as.factor(Model), frame=Subsample_Size,fill=factor(Model))) + 
  geom_histogram(binwidth = 0.01,alpha=0.8)+
  facet_grid(Type~Theta,scales = "free_x",shrink = TRUE)+
  geom_vline(data= Bias_Vline,aes(xintercept = Values,color="red"),linetype='dashed',size=1)+
  scale_color_discrete(name = "Bias values", labels = c("Real"))+
  labs(title=paste0("Real Model :",params$Model_Path," Data : Estimated Bias"),
      subtitle = "Sub-sample Size: {current_frame}")+
  transition_manual(Subsample_Size)+theme_light()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  ylab("Frequency")+xlab("Estimates")->Bias_1

anim_save(filename="Bias_1.gif",animation = Bias_1,
          path=here("Identical_r0","htmloutputs",params$Model_Path,"OSMAC","Plots"))

include_graphics2(here("Identical_r0","htmloutputs",params$Model_Path,"OSMAC","Plots","Bias_1.gif"))
```

## Utility values

```{r Animated Utility plots RM,fig.width=7,fig.height=7}
Utility %>%
  ggplot(.,aes(x=log(A_optimality),fill=Model))+
  geom_histogram()+facet_grid(~factor(Type))+
  theme_light()+ylab("Frequency")+xlab("log(A-optimality)")+
  transition_states(Subsample_Size)+
  ggtitle("Real Model : Subsample : {closest_state}")->U_1

anim_save(filename="U_1.gif",animation = U_1,
          path=here("Identical_r0","htmloutputs",params$Model_Path,"OSMAC","Plots"))

include_graphics2(here("Identical_r0","htmloutputs",params$Model_Path,"OSMAC","Plots","U_1.gif"))

Utility %>%
  ggplot(.,aes(x=log(D_optimality),fill=Model))+
  geom_histogram()+facet_grid(~factor(Type))+
  theme_light()+ylab("Frequency")+xlab("log(D-optimality) ")+
  transition_states(Subsample_Size)+
  ggtitle("Real Model : Subsample : {closest_state}")->U_2

anim_save(filename="U_2.gif",animation = U_2,
          path=here("Identical_r0","htmloutputs",params$Model_Path,"OSMAC","Plots"))

include_graphics2(here("Identical_r0","htmloutputs",params$Model_Path,"OSMAC","Plots","U_2.gif"))
```

# Comparison for Efficiency

```{r Comparison of Summary Data}
Final_MSE<-function(Real,Estimated_1,Estimated_2,Estimated_3,Estimated_4,Estimated_5,Model)
{
  For_Theta0<-sum((Real[1]-Estimated_1)^2,na.rm = TRUE)
  For_Theta1<-sum((Real[2]-Estimated_2)^2,na.rm = TRUE)
  For_Theta2<-sum((Real[3]-Estimated_3)^2,na.rm = TRUE)

  if(Model=="Model_1")
  {
    return((For_Theta0+For_Theta1+For_Theta2)/length(Estimated_1))
  }  
  if(Model=="Model_2" || Model=="Model_3")
  {
    For_Theta3<-sum((Real[4]-Estimated_4)^2,na.rm = TRUE)
    return((For_Theta0+For_Theta1+For_Theta2+For_Theta3)/length(Estimated_1))
  }
  if(Model=="Model_4")
  {
    For_Theta3<-sum((Real[4]-Estimated_4)^2,na.rm = TRUE)
    For_Theta4<-sum((Real[5]-Estimated_5)^2,na.rm = TRUE)
    return((For_Theta0+For_Theta1+For_Theta2+For_Theta3+For_Theta4)/length(Estimated_1))
  }
}

# Summarise A- optimality
Utility %>%
  dplyr::select(Type,Model,Subsample_Size,A_optimality) %>%
  group_by(Type,Model,Subsample_Size) %>%
  mutate(mean_Variance=mean(A_optimality))%>%
  distinct(mean_Variance) %>% 
  forestmangr::round_df(8)->All_VAR

# Summarise D- optimality
Utility %>%
  dplyr::select(Type,Model,Subsample_Size,D_optimality) %>%
  group_by(Type,Model,Subsample_Size) %>%
  mutate(mean_Info=mean(D_optimality))%>%
  distinct(mean_Info) %>% 
  forestmangr::round_df(8)->All_Info

if(params$Model_Path=="Model_1")
{
  # Summarize SMSE
  Est_Parameter %>%
        dplyr::select(Type,Model,Subsample_Size,Theta0,Theta1,Theta2) %>%
        group_by(Type,Model,Subsample_Size) %>%
        mutate( MSE=Final_MSE(Real = Real_Vline$Values,
                              Estimated_1=Theta0, 
                              Estimated_2=Theta1,
                              Estimated_3=Theta2,
                              Model = params$Model_Path ))%>%
        distinct(MSE) %>% 
        forestmangr::round_df(8)->All_MSE

  # Summarize SMSE MLE
  Est_Parameter %>%
        dplyr::select(Type,Model,Subsample_Size,Theta0,Theta1,Theta2) %>%
        group_by(Type,Model,Subsample_Size) %>%
        mutate( MSE=Final_MSE(Real = Estimated_Vline$Values,
                              Estimated_1=Theta0, 
                              Estimated_2=Theta1,
                              Estimated_3=Theta2,
                              Model = params$Model_Path ))%>%
        distinct(MSE) %>% 
        forestmangr::round_df(8)->All_MSE_MLE

  # Summarise Bias
  Est_Bias %>%
    dplyr::select(Type,Model,Subsample_Size,Theta0,Theta1,Theta2) %>%
    group_by(Type,Model,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + Theta2*Theta2)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias

  # Summarise Bias = sqrt(Bias^2)
  Est_Bias %>%
    dplyr::select(Type,Model,Subsample_Size,Theta0,Theta1,Theta2) %>%
    group_by(Type,Model,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + Theta2*Theta2) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared
}

if(params$Model_Path=="Model_2"||params$Model_Path=="Model_3")
{
  # Summarize SMSE
  Est_Parameter %>%
        dplyr::select(Type,Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
        group_by(Type,Model,Subsample_Size) %>%
        mutate( MSE=Final_MSE(Real = Real_Vline$Values,
                              Estimated_1=Theta0, 
                              Estimated_2=Theta1,
                              Estimated_3=Theta2,
                              Estimated_4=Theta3,
                              Model = params$Model_Path ))%>%
        distinct(MSE) %>% 
        forestmangr::round_df(8)->All_MSE

  # Summarize SMSE MLE
  Est_Parameter %>%
        dplyr::select(Type,Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
        group_by(Type,Model,Subsample_Size) %>%
        mutate( MSE=Final_MSE(Real = Estimated_Vline$Values,
                              Estimated_1=Theta0, 
                              Estimated_2=Theta1,
                              Estimated_3=Theta2,
                              Estimated_4=Theta3,
                              Model = params$Model_Path ))%>%
        distinct(MSE) %>% 
        forestmangr::round_df(8)->All_MSE_MLE

  # Summarise Bias
  Est_Bias %>%
    dplyr::select(Type,Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Model,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2 + Theta3*Theta3)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias

  # Summarise Bias = sqrt(Bias^2)
  Est_Bias %>%
    dplyr::select(Type,Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3) %>%
    group_by(Type,Model,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                            Theta2*Theta2 + Theta3*Theta3 ) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared
}

if(params$Model_Path=="Model_4")
{
  # Summarize SMSE
  Est_Parameter %>%
        dplyr::select(Type,Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
        group_by(Type,Model,Subsample_Size) %>%
        mutate( MSE=Final_MSE(Real = Real_Vline$Values,
                              Estimated_1=Theta0, 
                              Estimated_2=Theta1,
                              Estimated_3=Theta2,
                              Estimated_4=Theta3,
                              Estimated_5=Theta4,
                              Model = params$Model_Path ))%>%
        distinct(MSE) %>% 
        forestmangr::round_df(8)->All_MSE

  # Summarize SMSE
  Est_Parameter %>%
        dplyr::select(Type,Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
        group_by(Type,Model,Subsample_Size) %>%
        mutate( MSE=Final_MSE(Real = Estimated_Vline$Values,
                              Estimated_1=Theta0, 
                              Estimated_2=Theta1,
                              Estimated_3=Theta2,
                              Estimated_4=Theta3,
                              Estimated_5=Theta4,
                              Model = params$Model_Path ))%>%
        distinct(MSE) %>% 
        forestmangr::round_df(8)->All_MSE_MLE

  # Summarise Bias
  Est_Bias %>%
    dplyr::select(Type,Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
    group_by(Type,Model,Subsample_Size) %>%
    mutate(mean_Bias=mean(sqrt(Theta0*Theta0 + Theta1*Theta1 + 
                                 Theta2*Theta2 + Theta3*Theta3 + Theta4*Theta4)) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(8)->All_Bias

  # Summarise Bias = sqrt(Bias^2)
  Est_Bias %>%
    dplyr::select(Type,Model,Subsample_Size,Theta0,Theta1,Theta2,Theta3,Theta4) %>%
    group_by(Type,Model,Subsample_Size) %>%
    mutate(mean_Bias=mean(Theta0*Theta0 + Theta1*Theta1 + 
                            Theta2*Theta2 + Theta3*Theta3 + Theta4*Theta4) )%>%
    distinct(mean_Bias) %>% 
    forestmangr::round_df(4)->All_Bias_squared
}

# Summarise EMSE
inner_join(All_VAR,All_Bias_squared,by=c("Type","Model","Subsample_Size")) %>%
  mutate(EMSE=mean_Variance+mean_Bias) %>%
  select(Type,Model,Subsample_Size,EMSE)->All_EMSE
colnames(All_EMSE)<-c("Type","Model","Subsample_Size","EMSE")

# Full Data model Estimates
Full_Est_Parameter<-as.data.frame(t(Estimated_Vline$Values))
Full_Est_Bias<-as.data.frame(t(Cordeiro(XData = Simulated_Data[[2]]$Real_Model[,-1],
                                      With_bias = as.matrix(Full_Est_Parameter))))
colnames(Full_Est_Parameter)<-colnames(Full_Est_Bias)<-c(paste0("Theta",0:length(Simulated_Data$Basic$Theta[-1])))

Full_Bias<-sqrt(sum(Full_Est_Bias^2))
Temp_P<-invlogit(Simulated_Data[[2]]$Real_Model[,-1]%*% t(as.matrix(Full_Est_Parameter)))
Temp_W<-diag(as.vector(Temp_P*(1-Temp_P)))
Temp_Info<-t(Simulated_Data[[2]]$Real_Model[,-1])%*% Temp_W %*% Simulated_Data[[2]]$Real_Model[,-1]
Full_Variance<-tr(solve(Temp_Info))
Full_Information<-det(Temp_Info)

if(params$Model_Path=="Model_1")
{
  Full_SMSE<-Final_MSE(Real=Real_Vline$Values,
                Estimated_1 = Full_Est_Parameter$Theta0,
                Estimated_2 = Full_Est_Parameter$Theta1,
                Estimated_3 = Full_Est_Parameter$Theta2,
                Model=params$Model_Path)
}

if(params$Model_Path=="Model_2"||params$Model_Path=="Model_3")
{
  Full_SMSE<-Final_MSE(Real=Real_Vline$Values,
                Estimated_1 = Full_Est_Parameter$Theta0,
                Estimated_2 = Full_Est_Parameter$Theta1,
                Estimated_3 = Full_Est_Parameter$Theta2,
                Estimated_4 = Full_Est_Parameter$Theta3,
                Model=params$Model_Path)
}

if(params$Model_Path=="Model_4")
{
  Full_SMSE<-Final_MSE(Real=Real_Vline$Values,
                Estimated_1 = Full_Est_Parameter$Theta0,
                Estimated_2 = Full_Est_Parameter$Theta1,
                Estimated_3 = Full_Est_Parameter$Theta2,
                Estimated_4 = Full_Est_Parameter$Theta3,
                Estimated_5 = Full_Est_Parameter$Theta4,
                Model=params$Model_Path)
}

Full_EMSE<-Full_Variance+Full_Bias*Full_Bias

Full_Data<-as.matrix(Simulated_Data$All_Data$Real_Model)
Parameter_Data<-as.matrix(Est_Parameter[,-c(1:3)])
Parameter<-Parameter_Data[1,]

Predict_Logistic_Regression<-function(Parameter)
{
  Estimates<-matrix(Parameter)
  fitted_pi<-LaplacesDemon::invlogit(Full_Data[,-1]%*%Estimates)
  fitted_response<-ifelse(fitted_pi > 0.5,1,0)
  return(1-mean(fitted_response != Full_Data[,1]))
}

Accuracy<-apply(Parameter_Data,1,Predict_Logistic_Regression)
Accuracy_Data<-cbind.data.frame(Est_Parameter[,1:3],Accuracy)
Accuracy_Data<-Accuracy_Data %>% 
  group_by(Model,Type,Subsample_Size) %>%
  summarize(Average_Accuracy=mean(Accuracy))

remove(Temp_P,Temp_W,Temp_Info,Full_Data,Parameter_Data,Parameter,Accuracy)

# Store the Results
save(All_MSE,All_MSE_MLE,All_EMSE,All_Bias,All_VAR,All_Info,Accuracy_Data,
     file = here("Identical_r0","Summary",params$Model_Path,"Data","Efficiency",
                 "No_Correlated_Covariate_OSMAC.RData"))
```

## SMSE, EMSE, A-optimality, D-optimality, Bias Comparing Real Model and Assumed Models

```{r Final plots 1,fig.height=7,fig.width=20}
Plot_List<-list()

All_MSE %>%
  #ungroup(Type,Model,Subsample_Size) %>%
ggplot(.,aes(x=Subsample_Size,y=MSE,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("SMSE")+ggtitle("SMSE Comparison")->Plot_List[[1]]

All_MSE_MLE %>%
  #ungroup(Type,Model,Subsample_Size) %>%
ggplot(.,aes(x=Subsample_Size,y=MSE,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("SMSE")+ggtitle("SMSE Comparison for MLE")->Plot_List[[2]]

All_VAR %>%
  ungroup(Type,Model,Subsample_Size) %>%
ggplot(.,aes(x=Subsample_Size,y=mean_Variance,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("Mean Variance")+ggtitle("Estimated Variance Comparison")->Plot_List[[3]]

All_Info %>%
  ungroup(Type,Model,Subsample_Size) %>%
ggplot(.,aes(x=Subsample_Size,y=mean_Info,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("Average Information")+ggtitle("Estimated Model Information Comparison")->Plot_List[[4]]

All_Bias %>%
  ungroup(Type,Model,Subsample_Size) %>%
ggplot(.,aes(x=Subsample_Size,y=mean_Bias,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("Average Bias")+ggtitle("Estimated Bias Comparison")->Plot_List[[5]]

All_EMSE %>%
  ungroup(Type,Model,Subsample_Size) %>%
ggplot(.,aes(x=Subsample_Size,y=EMSE,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("EMSE")+ggtitle("EMSE Comparison")->Plot_List[[6]]

Accuracy_Data %>% 
  ggplot(.,aes(x=Subsample_Size,y=Average_Accuracy,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("Average Accuracy")+ggtitle("Average Accuracy")->Plot_List[[7]]

ggarrange(plotlist = Plot_List,common.legend = TRUE,nrow=1)

```

## SMSE, EMSE, A-optimality, D-optimality, Bias Comparing Real Model and Assumed Models with Full Data Model

```{r Final plots 2,fig.height=7,fig.width=25}
Plot_List<-list()

All_MSE %>%
  ungroup(Type,Model,Subsample_Size) %>%
  mutate(Efficiency=MSE/Full_SMSE) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("Relative Efficiency")+ggtitle("SMSE Efficiency")->Plot_List[[1]]

All_MSE_MLE %>%
  ungroup(Type,Model,Subsample_Size) %>%
  mutate(Efficiency=MSE/Full_SMSE) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("Relative Efficiency")+ggtitle("SMSE Efficiency for MLE")->Plot_List[[2]]

All_VAR %>%
  ungroup(Type,Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Variance/Full_Variance) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("Relative Efficiency")+ggtitle("Estimated Variance Efficiency")->Plot_List[[3]]

All_Info %>%
  ungroup(Type,Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Info/Full_Information) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("Relative Efficiency")+ggtitle("Estimated Model Information Efficiency")->Plot_List[[4]]

All_Bias %>%
  ungroup(Type,Model,Subsample_Size) %>%
  mutate(Efficiency=mean_Bias/Full_Bias) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("Relative Efficiency")+ggtitle("Estimated Bias Efficiency")->Plot_List[[5]]

All_EMSE %>%
  ungroup(Type,Model,Subsample_Size) %>%
  mutate(Efficiency=EMSE/Full_EMSE) %>%
ggplot(.,aes(x=Subsample_Size,y=Efficiency,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("Relative Efficiency")+ggtitle("EMSE Efficiency")->Plot_List[[6]]

Full_Data<-as.matrix(Simulated_Data$All_Data$Real_Model)
Parameter_Data<-as.matrix(Est_Parameter[,-c(1:3)])
Parameter<-Parameter_Data[1,]

fitted_pi<-as.vector(predict.glm(Full_model,newdata = as.data.frame(Simulated_Data$All_Data$Real_Model[,-1])))
fitted_response<-ifelse(fitted_pi > 0.5,1,0)
Full_Accuracy<-1-mean(fitted_response != Simulated_Data$All_Data$Real_Model[,1])

Accuracy_Data %>% 
  mutate(Relative_Accuracy=Average_Accuracy/Full_Accuracy) %>%
  ggplot(.,aes(x=Subsample_Size,y=Relative_Accuracy,color=Model))+geom_point()+
  geom_line()+scale_colour_viridis_d()+
  theme_light()+xlab("Sub sample size")+facet_wrap(~Type,ncol = 1)+
  ylab("Relative Average Accuracy")+ggtitle("Average Accuracy")->Plot_List[[7]]

ggarrange(plotlist = Plot_List,common.legend = TRUE,nrow=1)
```